Beyond Adaptive Web Design

### 2014-03-16 paper/paper.tex
fix: remove contractions

The basic approach how to provide an optimal viewing experience is to use ``Responsive web design'' which comes with fluid grids, flexible images and CSS3 media queries~\cite{responsive}. While creating flexible layouts is important, there's a lot more that requires remarkable adaptive web experiences. The principles of Adaptive Design are: ubiquity, flexibility, performance, enhancement and future-friendliness~\cite{adaptivesxsw}.

The power of the web is its ubiquity. No one knows what the landscape is going to appear in a couple years, but there's a reasonable probability the new devices in a few years from now will have access to the web. Because of this, we need to preserve and embrace the web's ubiquity. This requires us to provide full web experiences regardless of how and where people access the web. It's important to remain in creating flexible interfaces that can adapt to any screen size. Performance is often less important to everything else. ``Progressive enhancement'', feature detection and many other techniques allow us to increase experience that allow us to support more devices while still optimizing for the finest. The key aspect of ``Future Friendly'' thinking is to acknowledge and embrace unpredictability. Nobody knows where things are going but the one thing we can trust is the evolution.

Adaptive web design is fundamentally progressive enhancement, but it's being applied to a much larger and more diverse landscape. We now have Web-enabled smartphones, tablets, e-readers, netbooks, watches, TVs, phablets, notebooks, game consoles, cars and more. We also have many types of internet networks with different speed, latency and quality. Responsive design is also one technique in an adaptive web design strategy. Creating flexible layouts is important, but there are many more factors we need to think about. It's also important to consider as well ergonomics, touch capability, other input methods, internet connections and many other features that can be detected.

@@

Web browsers on TVs are actually really good but the input is terrible so no one uses them. Input is much more important to interface design than screen size. The input defines what a design needs to do in order to accomplish a task~\cite{adaptiveinput}. We also have wearable computers. The current Google Glass interface is limited to a floating display above head. We don't have full range of motion and, as a result, no spatial cues or depth in the interface.

@@

The optimal touch target size is 7mm, based on the average size of human finger tips and pads. CSS2.1 defines a pixel as 1/96 of an inch. So 7mm should be 30pixels~\cite{designingfortouch}. However, things aren't so easy because of dynamic viewports. It means that input elemtents in website have to be usually bigger.

@@

We can now use speech recognition to take a full control of the web site as we have access to continuous recognition stream. We don't need another input methods at all. A speech feedback for user interaction is also important and we can use speech synthesis feature for it.

@@

From controlling the user interface based on three dimensional motion as input, to combining device motion with another sensors as are location detection or video cameras~\cite{AccelerationAndGyroscope}. There's a lot of interesting interface designs made possible by Device Motion. It depends on us how we use them.

@@

We can detect user motion and use it as an alternatove part of input method. We can build web games and another real time interfaces. There's again a lot of interesting interface designs made possible.

@@

We all should be interested in page speed and mobile performance because people need fast experiences on the Web. There's lot of studies that highlight the importance of speed. More than half of people with a bad loading experience on mobile, won't come back~\cite{performancebrowsernetworking}. 73\% of mobile internet users say they've encountered Web pages that are too slow, a 1 second delay can result in a 7\% reduction in conversions~\cite{pagespeed}. Speed is an esencial part of our strategy and we should care about it. We can quantify its impact on the metrics and track it.

There are many different types of internet connections. Mobile network speeds have rose but this doesn't help much as page load times are still high as bandwidth increases. Latency is a much bigger fragment to download times. LTE standart reduces tower latency by several milliseconds but we still need to make a number of connections to download data to our devices~\cite{performancebrowsernetworking}.

@@

However some problems have been detected. The biggest one is that browser vendors don't use specification correctly and use own orientation ranges and directions. This causes some scroll issues.

@@

There are many commonly used web services that produce unnecessary traffic. Elements like videos or maps are a common part of the most web sites. All of them produces many requests and the web traffic so the website user doesn't want to watch the video or browse the map. There are also native mobile application for that services with better perfomance which provide full user experience. So what can we to do? This is a great opportunity to utilize conditional loading to serve the best experience for the right context.

We use the Mobile First principe so we make a default element static and pull in a static image. This approach saves a lot of web traffic and becasue of limited mobile data plans also user's money. We can then detect when it's appropriate to download the embedded map or to launch the native mobile application.

@@

By default we set an anchor to the location of the service website. When a user taps the component, the browser will try to open the native application where they can have a better experience. If the native experience is unavailable, the user is redirected to the service website, where they receive a full-screen experience. However, we don't always want to open the native application. We can download an embedded map or play the video directly after the tap. The user can interact with them directly in the web application so developers have to make a decision between these choices.

>>>>

The basic approach how to provide an optimal viewing experience is to use ``Responsive web design'' which comes with fluid grids, flexible images and CSS3 media queries~\cite{responsive}. While creating flexible layouts is important, there is a lot more that requires remarkable adaptive web experiences. The principles of Adaptive Design are: ubiquity, flexibility, performance, enhancement and future-friendliness~\cite{adaptivesxsw}.

The power of the web is its ubiquity. No one knows what the landscape is going to appear in a couple years, but there is a reasonable probability the new devices in a few years from now will have access to the web. Because of this, we need to preserve and embrace the web's ubiquity. This requires us to provide full web experiences regardless of how and where people access the web. It is important to remain in creating flexible interfaces that can adapt to any screen size. Performance is often less important to everything else. ``Progressive enhancement'', feature detection and many other techniques allow us to increase experience that allow us to support more devices while still optimizing for the finest. The key aspect of ``Future Friendly'' thinking is to acknowledge and embrace unpredictability. Nobody knows where things are going but the one thing we can trust is the evolution.

Adaptive web design is fundamentally progressive enhancement, but it is being applied to a much larger and more diverse landscape. We now have Web-enabled smartphones, tablets, e-readers, netbooks, watches, TVs, phablets, notebooks, game consoles, cars and more. We also have many types of internet networks with different speed, latency and quality. Responsive design is also one technique in an adaptive web design strategy. Creating flexible layouts is important, but there are many more factors we need to think about. It is also important to consider as well ergonomics, touch capability, other input methods, internet connections and many other features that can be detected.

@@

Web browsers on TVs are actually really good but the input is terrible so no one uses them. Input is much more important to interface design than screen size. The input defines what a design needs to do in order to accomplish a task~\cite{adaptiveinput}. We also have wearable computers. The current Google Glass interface is limited to a floating display above head. We do not have full range of motion and, as a result, no spatial cues or depth in the interface.

@@

The optimal touch target size is 7mm, based on the average size of human finger tips and pads. CSS2.1 defines a pixel as 1/96 of an inch. So 7mm should be 30pixels~\cite{designingfortouch}. However, things are not so easy because of dynamic viewports. It means that input elemtents in website have to be usually bigger.

@@

We can now use speech recognition to take a full control of the web site as we have access to continuous recognition stream. We do not need another input methods at all. A speech feedback for user interaction is also important and we can use speech synthesis feature for it.

@@

From controlling the user interface based on three dimensional motion as input, to combining device motion with another sensors as are location detection or video cameras~\cite{AccelerationAndGyroscope}. There is a lot of interesting interface designs made possible by Device Motion. It depends on us how we use them.

@@

We can detect user motion and use it as an alternatove part of input method. We can build web games and another real time interfaces. There is again a lot of interesting interface designs made possible.

@@

We all should be interested in page speed and mobile performance because people need fast experiences on the Web. There is lot of studies that highlight the importance of speed. More than half of people with a bad loading experience on mobile, will not come back~\cite{performancebrowsernetworking}. 73\% of mobile internet users say they have encountered Web pages that are too slow, a 1 second delay can result in a 7\% reduction in conversions~\cite{pagespeed}. Speed is an esencial part of our strategy and we should care about it. We can quantify its impact on the metrics and track it.

There are many different types of internet connections. Mobile network speeds have rose but this does not help much as page load times are still high as bandwidth increases. Latency is a much bigger fragment to download times. LTE standart reduces tower latency by several milliseconds but we still need to make a number of connections to download data to our devices~\cite{performancebrowsernetworking}.

@@

However some problems have been detected. The biggest one is that browser vendors do not use specification correctly and use own orientation ranges and directions. This causes some scroll issues.

@@

There are many commonly used web services that produce unnecessary traffic. Elements like videos or maps are a common part of the most web sites. All of them produces many requests and the web traffic so the website user does not want to watch the video or browse the map. There are also native mobile application for that services with better perfomance which provide full user experience. So what can we to do? This is a great opportunity to utilize conditional loading to serve the best experience for the right context.

We use the Mobile First principe so we make a default element static and pull in a static image. This approach saves a lot of web traffic and becasue of limited mobile data plans also user's money. We can then detect when it is appropriate to download the embedded map or to launch the native mobile application.

@@

By default we set an anchor to the location of the service website. When a user taps the component, the browser will try to open the native application where they can have a better experience. If the native experience is unavailable, the user is redirected to the service website, where they receive a full-screen experience. However, we do not always want to open the native application. We can download an embedded map or play the video directly after the tap. The user can interact with them directly in the web application so developers have to make a decision between these choices.
